
% Chapter Introduction

% What would be included in this chapter?
This chapter provides an introduction about the motivation of this thesis in section \ref{section:motivation}. Moreover, an outline is listed in section \ref{section:thesis outline}.

% Section Motivation

\section{Motivation} \label{section:motivation}
% Short description of what is Linked Data, Question Answering, SPARQL. What is the relationship between these concepts? What is the current situation and what are the current demands? What is the (goal) task of this thesis? Why is it important to realize this goal? why is it useful to translate natural language to SPARQL? Who can benefit from it and how can they benefit from it?
The World Wide Web is quickly evolving nowadays and has now become a huge network contaning various kinds of resources for billions of users to interact with. A majority of the documents on the web are formatted texts in Hypertext Markup Language (HTML), serving as the content that can be rendered on computer browsers for humans to read and understand. In order to find a resource satisfying specific needs, a web user normally needs to rely on the help of search engines to retrieve and filter results from innumerable documents on the web. Even so, it takes quite more time for humans to distinguish useful materials from others than machines that can scan large number of files almost simultaneously. However, finishing such kind of tasks requires the capability of understanding the documents to be scanned and the traditional file format like HTML has made it difficult for machines to do so.

The Semantic Web is the concept of a Web where data and information can be manipulated by machines automatically \cite{Shadbolt2006}. There has been a set of standards for altering the current web to be more machine-readable and processable for automatic machine agents. In order to help achieve the potential of the current web, a series of relevant technologies including mainly Resource Description Framework (RDF) \cite{Cyganiak2014} and Ontology have been introduced. With the help of these tools, an increasing number of documents containing uniform organized data have been published conveniently on the web. One notable example of this is cross-domain Linked Datasets such as DBpedia \cite{Auer2007}. DBpedia contains RDF documents that represents knowledge information extracted from Wikipedia websites, and all the documents with links to other datasets on the web consititute an interlinked ontology model. DBpedia also provides an open API for users to ask complicated queries against those documents.

SPARQL is a language designed for humans to query and manipulate the information sources contained in an RDF store or online RDF graph content, and is by far the recommended standard \cite{Harris2013}. SPARQL has already been widely supported by large open datasets like DBpedia. Though the data queried by SPARQL is made for publicity and openness, the use of it has yet been spreaded out of a group of experts with prior knowledge specific to some certain domain. For example, if a user wants to ask for a list of books belonging to some certain category, he or she needs to have prior knowledge about the concepts and relations involved in describing books in RDF. The root of this problem is the gap between the natural language used by non-experts and the query language that consists of unique syntax, semantic and domain-specific vocabulary. 

The motivation of this thesis is to bridge this gap of natural language and SPARQL. While the natural language and SPARQL are both able to be represented as sequences of pre-defined tokens, this can be categorized as a translation problem. In rencent years, the application of neural networks in machine translation has achieved great improvement on translation results than previously applied statistical and phrase-based methods \cite{Moussallem2017}. Therefore, we think of altering neural network models that have proven great performance in machine translation to apply on the task of translating natural language to the expressions written in SPARQL. We look into the effects of such alteration by conducting several experiments, and making parallel comparisons between different models with varying network configurations. 

\section{Thesis Outline} \label{section:thesis outline}
% How many chapters does this thesis have? What are the main contents that would be described in each chapter? 

Chapter \ref{chapter:background} presents the notion of Semantic Web and its corresponding technologies, research in the subject of neural machine translation under the area of deep learning, and the past work closely related to this thesis. 

(TBD) Chapter \ref{chapter:methodology} describes the research method used in this thesis.

Chapter \ref{chapter:experiments} shows the experiments carried out to investigate better neural network models on the task of translating natural language to SPARQL, the datasets applied, and the corresponding results in textual and tablular form.

Chapter \ref{chapter:analysis} depicts the analysis derived from the experiment results exhibited in chapter \ref{chapter:experiments}.

Chapter \ref{chapter:conclusion} brings a summary in general and provides an outlook for the future work.